Convergence rate of MAP estimates \\ for the exponential family

### 2021-10-21 aistats.tex
Add gaussian variance MLE

\subsection{Bias of a Gaussian Variance MLE}
\rlp{prove that the bias decreases like $1/n^2$ for gaussian variance and gaussian mean.}

@@

We use this section to give more details on the (stochastic) mirror descent algorithm.
One perspective on gradient descent with step-sizes, 
the update $\theta_{n+1} = \theta_n - \gamma \nabla f(\theta_n)$ is to view it 
as the minimization of the linear approximation of $f$ at $\theta_n$, 
$f(\theta) \approx f(\theta_n) + \lin{\nabla f(\theta_n), \theta - \theta_n}$
and a quadratic penalty scaled by ${1}/{\gamma}$, 
\alignn{
	\theta_{n+1} = \arg\min_\theta f(\theta_n) + \lin{\nabla f(\theta_n), \theta - \theta_n} + \frac{1}{\gamma} \frac{1}{2}\norm{\theta - \theta_n}^2.
}
Mirror descent generalizes the above, using the Bregman divergence induced by a (Legendre) function $A$ instead of the Euclidean norm, 
\alignn{\label{eq:mirror-descent-primal}
	\theta_{n+1} = \arg\min_\theta f(\theta_n) + \lin{\nabla f(\theta_n), \theta - \theta_n} + \frac{1}{\gamma} {\theta - \theta_n}^2.
}
Mirror descent coincides with gradient descent if $A(\theta) = \frac{1}{2}\norm{\theta}^2$.
As \cref{eq:mirror-descent-primal} is convex, the minimum is at a stationary point, 
found by taking the derivative and setting to 0, 
leading to the update $\theta_{n+1}$ satisfying
\alignn{
	\nabla f(\theta_n) + \frac{1}{\gamma}\paren{\nabla A(\theta_{n+1}) - \nabla A(\theta_n)} = 0
	&&\implies&&
	\nabla A(\theta_{n+1}) = \nabla A(\theta_n) - \gamma \nabla f(\theta_n).
}
Expressed with the dual parameters $\mu_{n} = \nabla A(\theta_{n})$, 
$\mu_{n+1} = \mu_n - \gamma \nabla f(\theta_n)$.

>>>>

\subsection{Bias of a Gaussian Variance MLE}
For a Gaussian variance model, the MLE follows a scaled $\chi^2(n)$ distribution. 
This means that
\alignn{
	\frac{\m^*}{\tilde \m_n} 
	= \frac{\tilde\nat_n}{\nat^*}
	=\expect{\frac{\hat\nat_n}{\nat^*}}
	=\expect{\frac{\m^*}{\hat \m_n} }
	=\expect{\frac{n}{\chi^2(n)}}
	= \frac{n}{n-2}\; .
}
Consequently
\alignn{
	\bregmanconj(\m^*, \tilde \m_n) 
	&= \half \paren{\frac{n}{n-2} -1 + \log\frac{n-2}{n}}\\
	&= \frac{1}{n-2} + \half \log\paren{1 - \frac{2}{n}}\\
	&\leq \inv{n-2} - \inv{n} 
	= \frac{2}{n(n-2)}
}
so the bias of the MLE of a Gaussian Variance decreases like $O\paren{\inv{n^2}}$.

@@

We use this section to give more details on the (stochastic) mirror descent algorithm.
One perspective on gradient descent with step-sizes, 
the update $\theta_{n+1} = \theta_n - \gamma \nabla f(\theta_n)$ is to view it 
as the minimization of the linear approximation of $f$ at $\theta_n$, 
$f(\theta) \approx f(\theta_n) + \lin{\nabla f(\theta_n), \theta - \theta_n}$
and a quadratic penalty scaled by ${1}/{\gamma}$, 
\alignn{
	\theta_{n+1} = \arg\min_\theta f(\theta_n) + \lin{\nabla f(\theta_n), \theta - \theta_n} + \frac{1}{\gamma} \frac{1}{2}\norm{\theta - \theta_n}^2.
}
Mirror descent generalizes the above, using the Bregman divergence induced by a (Legendre) function $A$ instead of the Euclidean norm, 
\alignn{\label{eq:mirror-descent-primal}
	\theta_{n+1} = \arg\min_\theta f(\theta_n) + \lin{\nabla f(\theta_n), \theta - \theta_n} + \frac{1}{\gamma} \bregman(\theta, \theta_n).
}
Mirror descent coincides with gradient descent if $A(\theta) = \frac{1}{2}\norm{\theta}^2$.
As \cref{eq:mirror-descent-primal} is convex, the minimum is at a stationary point, 
found by taking the derivative and setting to 0, 
leading to the update $\theta_{n+1}$ satisfying
\alignn{
	\nabla f(\theta_n) + \frac{1}{\gamma}\paren{\nabla A(\theta_{n+1}) - \nabla A(\theta_n)} = 0
	&&\implies&&
	\nabla A(\theta_{n+1}) = \nabla A(\theta_n) - \gamma \nabla f(\theta_n).
}
Expressed with the dual parameters $\mu_{n} = \nabla A(\theta_{n})$, 
$\mu_{n+1} = \mu_n - \gamma \nabla f(\theta_n)$.
